---
title: "Data prep for food hackathon"
author: "Catalina Moreno"
date: '`r format(Sys.time(), format = "%B %d, %Y")`'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE, 
                      message = FALSE,
                      comment = NA)
library(tidyverse)
library(googlesheets)
```

## Background

Prepare data for PFC hackathon according to the following data model:


```{r, data_model}
## read in data model from google sheets
# gs_auth(new_user = TRUE) #auth account (one time action)
data_mod <- gs_read(gs_title("fields_and_descriptions"))

dat_mod_col_names <- data_mod %>% pull(field)

dat_mod_col_names
```

**Note:** 

* added open_time1 and open_time2 etc to data model on Google Drive for sites that have morning and afternoon open times

* Recommend to wait to assign UID until all compiled, then check name and address -- this is unique key

* What is long term ETL? How will pull future data, from shared drive or putting into db?

```{r, prep_future_ready_tables}
## create empty dataframe according todatamodel
dat1 <- data.frame(matrix(ncol = length(dat_mod_col_names), nrow = 0))
colnames(dat1) <- dat_mod_col_names
```

## data-conveniencesupermarkets

```{r, convenience_1}
data_convenience <- gs_read(gs_title("data-conveniencesupermarkets"))
head(data_convenience)
```

```{r, convenience_2}
## map data_convenience to dat1
data_convenience <- dat1 %>% 
  bind_rows(data_convenience %>% 
              mutate(address = ifelse(is.na(`Street Name`), 
                                      `Street #`, paste(`Street #`, `Street Name`))) %>% 
              select(name = Name,
                     type = Category,
                     address, 
                     state = State, 
                     zip_code = Zip, 
                     latitude = Lat, 
                     longitude = Lon)) #%>%  ## don't know day of operation yet, leave all NA
  #mutate_at(vars(contains("day")), ~replace(., is.na(.), 0))

head(data_convenience)
## write out as .csv 
write_excel_csv(data_convenience, "data_convenience_cleaned.csv")
```

**Missing information on hours/days of operation as well as SNAP, WIC, FMNP, fresh_produce-healthy, MRFEI_score info.**

## Allegheny_County_Farmers_Markets_Locations_2017

```{r, agh_farmers_market_1}
agh_farm_markets <- gs_read(gs_title("Allegheny_County_Farmers_Markets_Locations_2017"))
head(agh_farm_markets)
```

From `Day_Time` variable, need to be able to extract weekday(s), time of day open, months open. Write functions to extract this info (note these could be improved and will require further testing/updates):

```{r, agh_farmers_market_2}
source("support_funs.R")
## process and augment weekdays for each row of dataset 
multi_day_info <- vector("list", length = nrow(agh_farm_markets))

for(i in 1:nrow(agh_farm_markets)) {
  multi_day_info[[i]] <- weekday_calc(agh_farm_markets$Day_Time[i]) %>% 
    mutate(FID = agh_farm_markets$FID[i]) ## for join later
}


multi_day_info <- bind_rows(multi_day_info)

## go wide, to match data model:
multi_day_info <- multi_day_info %>% select(weekday, FID) %>% mutate(value = 1) %>% tidyr::spread(key = "weekday", value = "value") %>% mutate_at(vars(contains("day")), ~replace(., is.na(.), 0))

```

```{r, agh_farmers_market_3}
## using open time, month functions and left join wide weekday above into one dataframe that fits data model:
agh_farm_markets <- dat1 %>% 
  bind_rows(agh_farm_markets %>% 
              left_join(multi_day_info, by = "FID") %>% 
              rowwise() %>% 
              mutate(type = "Farmers Market",
                     weekday_val = str_trim(str_extract(Day_Time, "^[:alpha:]* ")),
                     open_time1 = time_calc(Day_Time)$open_time,
                     close_time1 = time_calc(Day_Time)$close_time,
                     date_from = season_calc(Season)$date_from,
                     date_to = season_calc(Season)$date_to) %>% ## this may not be robust
              select(type, open_time1, close_time1, date_from, date_to,
                     Monday, Tuesday, Wednesday, Thursday, Friday, Saturday, Sunday,
                address = Address,
                     name = Name, ## ? is this the correct farmers market name??
                     city = City,
                     state = State, 
                     zip_code = Zip, 
                     location_description = Location__,
                     latitude = X, 
                     longitude = Y)) 

## check
head(agh_farm_markets)

## write out as .csv 
write_excel_csv(agh_farm_markets, "agh_farm_markets_cleaned.csv")
```

**Missing SNAP, WIC, FMNP, fresh_produce-healthy, MRFEI_score info.**

If Green Grocer -- process name such that "Green Grocer (Store Name Here)

## growpghgardens201712

```{r, readin3}
growpgh <- gs_read(gs_title("growpghgardens201712_readin"))
head(growpgh)

```

```{r, growpgh}
growpgh <- dat1 %>% bind_rows(growpgh %>% select(name = urban_grower, 
                                                 type = category,
                                                 address = street_address,
                                                 city,
                                                 state,
                                                 zip_code,
                                                 latitude, 
                                                 longitude
                                                 ))

## write out as .csv 
write_excel_csv(growpgh, "growpghgardens_cleaned.csv")
```


**check addresses (not necessarily a street address), want to filter city to == Pittsburgh? Missing open day/time, SNAP etc info**

## PA

```{r, readinpa}
PA <- gs_read(gs_title("PA"))
head(PA)
```

Check ADDRESS2

```{r, address2}
PA %>% filter(!is.na(ADDRESS2)) %>% select(ADDRESS, ADDRESS2)
PA %>% distinct(County)
```

```{r, makepa}
PA <- dat1 %>% 
  bind_rows(PA %>% filter(County %in% "ALLEGHENY") %>% 
              mutate(address = paste(ADDRESS, ADDRESS2)) %>% 
              select(name = STORE_NAME, 
                          longitude,
                          latitude, 
                          address, # check linking together is correct
                          city = CITY,
                          state = STATE, 
                          zip_code = ZIP5))

write_excel_csv(PA, "PA_ALLEGHENY_cleaned.csv")

```

**Check if should filter beyond Allegheny county; missing open day/time and SNAP etc info**
