---
title: "eda"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE, 
                      comment = NA)
```

```{r}
library(tidyverse)
```

## background

The purpose of this document is to understand current merged data set. Specifically: 

* how to handle duplicates

* other irregularities (read lat/long)

* how to show gaps in data missing

* how to gain insight from current version, for use in web app

## read in data from GitHub

```{r}
dat <- read_csv("https://raw.githubusercontent.com/cgmoreno/food-access-map-data/master/merged_datasets.csv")
```

```{r}
# View(dat %>% head)
names(dat)
```
name

```{r}
dat %>% select(name) %>% head()

dat %>% count(name) %>% filter( n > 1)
```

type
```{r}
dat %>% count(type)
```
```{r}
dat %>% 
  ggplot(mapping = aes(x = type)) + geom_bar(fill = "blue") + coord_flip()
```

text processing on NA
```{r}
dat %>% filter(type %in% NA)
dat %>% filter(type %in% NA) %>% mutate(name = str_to_upper(name)) %>% 
  filter(str_detect(name, "RITE AID"))
```

RITE AID as convenience store

```{r}
dat %>% filter(is.na(type))
```
aldi, cogo, cvs, walgreen getgo, dollar general, family dollar, unimart, getgo, target, 
kuhns, sheets, giant eagle, gordon food service, walmart, sunoco, costco, 7-Elevon, whole food, trader joes
```{r}
supermarket <- c("ALDI", "TARGET", "KUHN", "GIANT EAGLE", "GORDON FOOD SERVICE", "WALMART", "COSTCO", "WHOLE FOODS", "TRADER JOES")
convenience_store <- c("COGO", "CVS", "WALGREEN", "GETGO", "DOLLAR GENERAL", "FAMILY DOLLAR", "UNIMART", "SHEETZ", "SUNOCO", "7-ELEVEN", "RITE AID")
```

```{r}
dat %>% filter(is.na(type)) %>% nrow
```
```{r}
dat %>% filter(is.na(type)) %>% mutate(name = str_to_upper(name)) %>% 
  filter(str_detect(name, "ALDI|TARGET|KUHN|GIANT EAGLE|GORDON FOOD SERVICE|WALMART|COSTCO|WHOLE FOODS|TRADER JOES")) %>% nrow()# 157
```
```{r}
dat %>% filter(is.na(type)) %>% mutate(name = str_to_upper(name)) %>% 
  filter(str_detect(name, "COGO|CVS|WALGREEN|GETGO|DOLLAR GENERAL|FAMILY DOLLAR|UNIMART|SHEETZ|SUNOCO|7-ELEVEN|RITE AID")) %>% nrow()


```
~ half, modify

```{r}
dat <- dat %>% mutate(type = ifelse(str_detect(str_to_upper(name), "COGO|CVS|WALGREEN|GETGO|DOLLAR GENERAL|FAMILY DOLLAR|UNIMART|SHEETZ|SUNOCO|7-ELEVEN|RITE AID"), "Convenience Store", type),
                      type = ifelse(str_detect(str_to_upper(name), "ALDI|TARGET|KUHN|GIANT EAGLE|GORDON FOOD SERVICE|WALMART|COSTCO|WHOLE FOODS|TRADER JOE|SHOP N SAVE"), "Supermarket", type))
```

```{r}
dat %>% 
  ggplot(mapping = aes(x = type)) + geom_bar(fill = "blue") + coord_flip()
```

500 remaining, unknown how to classify name (not known chain) or type. 

address
```{r}

```

zip_code, city, state

lat/long

date/times day of week

attributes SNAP, WIC, FMNP, fresh_produce_healthy, MRFEI_score, food_bucks, 
```{r}
dat %>% 
  select(SNAP, WIC, FMNP, `fresh_produce-healthy`, MRFEI_score, food_bucks, type) %>% 
  tidyr::gather(key = "key", value = "value", -type) %>% 
  ggplot(mapping = aes(x = value, fill = type)) + geom_bar() + facet_wrap(~key, scales = "free")
```

geocode info geoid vs ward vs hood 


More on type:

read in WPRDC sources:

```{r}
food_inspec <- read_csv("wprdc_food_inspec_for_type.csv")
food_inspec %>% count(description)
```

```{r}
food_inspec <- food_inspec %>% filter(description %in% c("Retail/Convenience Store", "Supermarket", "Chain Retail/Convenience Store", "Chain Supermarket"))
```

```{r}
food_inspec %>% head()
food_inspec %>% select(facility_name, description) %>% 
  group_by(facility_name) %>% summarise(description = first(description)) %>% 
  inner_join(dat %>% select(name, type), 
             by = c("facility_name" = "name")) %>% 
  filter(is.na(type))
```

bring in 96 types
rest will have to do somewhat manually. think of better way. 

```{r}
dat %>% filter(is.na(type)) %>% 
  mutate(new_name = str_to_upper(str_replace_all(name, " |[:punct:]|[:digit:]", ""))) %>% 
  select(id, name, type, new_name) %>% 
  count(new_name) %>% 
  filter(n > 10) %>% 
  ggplot(mapping = aes(x = new_name, y = n)) + geom_col() + coord_flip()
```
```{r}
dat %>% filter(is.na(type)) %>% nrow()

dat %>% filter(is.na(type)) %>% 
  mutate(new_name = str_to_upper(str_replace_all(name, " |[:punct:]|[:digit:]", ""))) %>% 
  select(id, name, type, new_name) %>% 
  count(new_name) %>% 
  filter(n > 10) %>%
  pull(n) %>% sum()
```

(capturing ~ 1/2 of the missing observations)

this represents


