---
title: "Fix Type Field in Merged Dataset"
author: "Catalina Moreno"
date: "November 12, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE, 
                      comment = NA)
```

```{r}
library(tidyverse)
```

## Background

The purpose of this document is to understand where the current version of the merged data set is missing the `type` field and populate that information when possible. Population could occur thru the WPRDC food inspection data set or using assumptions based on public knowledge regarding store brands. 

This is effort important because `type` affects how some rules for key labels are applied downstream (SNAP, WIC, etc.)


## Read in data from GitHub

The current version of the merged data set:
```{r}
dat <- read_csv("https://raw.githubusercontent.com/cgmoreno/food-access-map-data/master/merged_datasets.csv")
```

The WPRDC food inspection data, which contains `description` field analogous to `type` in our data set:
```{r}
food_inspec <- read_csv("wprdc_food_inspec_for_type.csv")
food_inspec <- food_inspec %>% filter(description %in% c("Retail/Convenience Store", "Supermarket", "Chain Retail/Convenience Store", "Chain Supermarket"))
```

## Address missing `type`

Check `type`:

```{r}
dat %>% count(type) %>% arrange(desc(n))
```

Note the large count of NA values. 

What are WPRDC types (`description` field):

```{r}
food_inspec %>% distinct(description)
```

Bring in WPRDC info to see if can populate type, changing to Convenience Store (remove Retail, collapse with Chain) and Supermarket (collapse Chain). 

```{r}
food_inspec <- food_inspec %>% mutate(description = str_replace(description, "Retail/", ""),
                       description = str_replace(description, "Chain ", "")) 
```

```{r}
food_inspec %>% select(facility_name, description) %>% 
  group_by(facility_name) %>% summarise(description = first(description)) %>% 
  inner_join(dat %>% select(name, type), 
             by = c("facility_name" = "name")) %>% 
  filter(is.na(type)) %>% nrow()
```

The WPRDC food inspec data populates 96 of the 1040 missing `types`.

As a validation step, check how WPRDC data set compares to those that have `type`, do `type` and description align? 

```{r}
food_inspec %>% select(facility_name, description) %>% 
  group_by(facility_name) %>% summarise(description = first(description)) %>% 
  inner_join(dat %>% select(name, type), 
             by = c("facility_name" = "name")) %>% 
  filter(!is.na(type), description == type) %>% nrow()
```

Do any not align? 

```{r}
food_inspec %>% select(facility_name, description) %>% 
  group_by(facility_name) %>% summarise(description = first(description)) %>% 
  inner_join(dat %>% select(name, type), 
             by = c("facility_name" = "name")) %>% 
  filter(!is.na(type), description != type) %>% nrow()
```

Yes, print these out:

```{r}
food_inspec %>% select(facility_name, description) %>% 
  group_by(facility_name) %>% summarise(description = first(description)) %>% 
  inner_join(dat %>% select(name, type), 
             by = c("facility_name" = "name")) %>% 
  filter(!is.na(type), description != type)
```

I would classify Aldi and Wal-Mart as Supermarket... check if any other Aldi or Wal-Mart in WPRDC data set:

```{r}
food_inspec %>% filter(str_detect(facility_name, "Aldi")) %>% count(description)
```
```{r}
food_inspec %>% filter(str_detect(facility_name, "Wal-Mart")) %>% count(description)

```

This is a limitation with the WPRDC data set. It has inconsistencies in `description` within the same store brand (though only 2 instances of this). For now, only bring in `description` where `type` is missing, and keep note of this limitation in case a better data set becomes available in the future. 

As next step, permanently join in `food_insepc` `description` information to populate those missing `type`.

```{r}
dat <- dat %>% left_join(food_inspec %>% select(facility_name, description),
                         by = c("name" = "facility_name")) %>% 
  mutate(type = ifelse(is.na(type), description, type)) %>% select(-description)
dat %>% count(type)
```

Note the new reduction in NA counts. 

For those remaining, next try text processing:

```{r}
dat %>% filter(type %in% NA) %>% mutate(name = str_to_upper(name),
                                        name = str_replace_all(name, "[:digit:]|[:punct:]", ""),
                                        name = str_trim(name)) %>% 
  count(name) %>% arrange(desc(n))
```

Via manual effort, build look up tables to assign `type` at these stores w/ > 1 count based on public domain knowledge:


```{r}
supermarket <- c("ALDI", "TARGET", "KUHN","KUHNS","KUHNS MARKET", "SAMS CLUB", "GIANT EAGLE", "GORDON FOOD SERVICE", "WALMART","WAL MART", "KMART", "COSTCO", "WHOLE FOODS", "TRADER JOES", "SAVALOT", "SAVEALOT")
supermarket <- tibble(name2 = supermarket) %>% mutate(type2 = "Supermarket")
# supermarket %>% count(name2) %>% filter(n > 1)
```


```{r}
# assume APLUS is sunoco, parmar is gas station too
convenience_store <- c("COGO","COGOS", "CVS", "CVS PHARMACY", "CVS PHARMACY INC", "WALGREEN","WALGREENS", "GETGO","GET GO", "DOLLAR GENERAL","DOLLAR GENERAL STORE", "FAMILY DOLLAR", "DOLLAR TREE", "SHEETZ", "SUNOCO", "SPEEDWAY", "7-ELEVEN", "RITE AID", "APLUS","APLUS H", "APLUS MINI MART", "ELEVEN H","ELEVEN STORE A", "ELEVEN STORE  A", "ELEVEN STORE  B", "ELEVEN STORE A A", "ELEVEN STORE B B", "ELEVEN STORE J J", "PAR MAR STORE", "UNI MART", "UNIMART", "CIRCLE K")
convenience_store <- tibble(name2 = convenience_store) %>% mutate(type2 = "Convenience Store")
# convenience_store %>% count(name2) %>% filter(n > 1)
```

Bring in this information 
```{r}
dat <- dat %>% mutate(name2 = str_to_upper(name),
               name2 = str_replace_all(name2, "[:digit:]|[:punct:]", ""),
               name2 = str_trim(name2),
               name2 = ifelse(is.na(type), name2, NA)) %>% # to only join to those missing
  left_join(convenience_store %>% bind_rows(supermarket), by = c("name2")) %>% 
  mutate(type = ifelse(is.na(type), type2, type)) %>% 
  select(-type2, -name2)
```

```{r}
dat %>% count(type)
```

Better. For now, will have to leave as is, as those remaining have 1 value for each name, and processing would require manually sifting through these ~ 400 values. Which data sources do these observations still missing `type` come from?

```{r}
dat %>% filter(is.na(type)) %>% count(source) %>% rename(count_missing_after = n) %>% 
  left_join(dat %>% count(source) %>% rename(count_orig = n), by = "source") %>% 
  select(source, count_orig, count_missing_after)
```

GPCFB Green Grocer and Just Harvest - Fresh Access Markets. 

Both of these should all be `type` = "Farmers Market" based on previously discussed rules with PFPC. 


```{r}
dat %>% filter(str_detect(name, "Green Grocer")) %>% select(name, type, source)
```



```{r}
dat %>% filter(str_detect(source, "Fresh Access Market")) %>% select(name, type, source)
```

These should both me modified during data prep. Apply now to see what remaining count of `type` = NA is:

```{r}
dat <- dat %>% mutate(type = ifelse(str_detect(name, "Green Grocer"), "Farmers Market", type),
                      type = ifelse(str_detect(source, "Fresh Access Market"), "Farmers Market", type))
```

```{r}
dat %>% count(type)
```

Final read out on sources and before vs. after this script the reduction in NA values for `type`:

```{r}
dat %>% filter(is.na(type)) %>% count(source) %>% rename(count_missing_after = n) %>% 
  left_join(dat %>% count(source) %>% rename(count_orig = n), by = "source") %>% 
  select(source, count_orig, count_missing_after)
```

## Summary

Through this document the following steps have been developed to handle missing `type` field:

* leverage WPRDC food inspection data set to bring in `description` field to current data set where `type` is missing 

* manual look up / make assignment based on public domain knowledge to brand name Supermarket and Convenience Stores

* fix Just Harvest - Fresh Access Markets and Green Grocer sources

This leaves ~ 400 missing `type` observations that can be improved upon. 

These steps should be incorporated in the data prep procedure (note, doing so will make new renderings of this document not informative, as it is linked directly to the most current merged data set on GitHub). 

## Appendix

Check names:

```{r}
names(dat)
```

